---
title: "basic_analysis"
output: html_document
date: "2026-04-22"
editor_options: 
  chunk_output_type: console
---

This is my setup code!
```{r}
# Libraries required for doc
library(knitr)
library(psych)
library(stats)
library(tidyverse)
library(dplyr)
library(ggplot2)

setwd('/Users/evaliss/Desktop/WAVElab_code/epCD/data/pilot 1')
```

Import my data and add necessary columns
```{r}
df <- read.csv("dataframe_answer.csv", header = TRUE)

## defining function
classify_effort <- function(firstPos, secondPos) {
  if (firstPos == secondPos) {
    return("No Change")
  } else if ((firstPos == "U" && secondPos == "C") || (firstPos == "C" && secondPos == "D")) {
    return("Less Effort")
  } else if ((firstPos == "C" && secondPos == "U") || (firstPos == "D" && secondPos == "C")) {
    return("More Effort")
  }
}

##creates and labels halves of trials
divide_halves <- function(trueTrial) {
  if (trueTrial <= (max(df$trial_number)/2)) {
    return("First Half")
  } else if (trueTrial > (max(df$trial_number)/2)){
    return("Second Half")
  }
}

##Takes out 6-trial participant currently. TO MODIFY --> disinclude anyone with less than 48 original trials.
grouped_participant_df <- df |> 
  group_by(participant_id) |> # first line group-by just tells you how you want to clump it together
  summarize(meanRT = mean(response_time), totalRT = n()) # this is where you tell it what new column you want to see

badParticipant_list <- grouped_participant_df[grouped_participant_df$totalRT < 48, ]$participant_id
# View(grouped_participant_df)

# calculates outlier range and narrows down grouped_participant_df
upper_pno <- quantile(grouped_participant_df$meanRT, 0.25) - sd_amount * IQR(grouped_participant_df$meanRT)
lower_pno <- quantile(grouped_participant_df$meanRT, 0.75) + sd_amount * IQR(grouped_participant_df$meanRT)

grouped_participant_df <- grouped_participant_df |>
  filter(meanRT < upper_bound) |>
  filter(meanRT > lower_bound)

# excludes bad participants with less than 48 trials + keeps good participants within outlier range
goodParticipant_list <- grouped_participant_df$participant_id
df <- df |>
  filter(!(participant_id %in% badParticipant_list)) |> 
  filter(participant_id %in% goodParticipant_list)


##exclude outlier trials
sd_amount <- 1.5
# "Extreme" outliers use a multiplier of 3.0 instead of 1.5
lower_bound <- quantile(df$response_time, 0.25) - sd_amount * IQR(df$response_time)
upper_bound <- quantile(df$response_time, 0.75) + sd_amount * IQR(df$response_time)

df <- df |>
  filter(response_time < upper_bound) |>
  filter(response_time > lower_bound)

## mutate (aka add column)
df <- df |> 
  mutate(effortCategory = mapply(classify_effort, firstCupPosition, secondCupPosition))

df <- df %>%
  mutate(handType = ifelse(grepl("Side", firstStim), "Side",
                    ifelse(grepl("Up", firstStim), "Up", "None")))

##runs various things
df <-  df |> 
  mutate(trialHalf = mapply(divide_halves, trial_number))

write.csv(df,"dataframe_answer_effort_malware.csv")

max(df$response_time)
min(df$response_time)
         
```


making a plot :D
```{r}
df <- read.csv("dataframe_answer_effort_malware.csv", header = TRUE)

class(df)
grouped_effortCategory <- df %>% 
  group_by(participant_id, effortCategory, handType, trialHalf) |> # first line group-by just tells you how you want to clump it together
  summarize(meanRT = mean(response_time), 
            meanAcc = mean(accuracy),
            .groups = "drop")
# View(grouped_effortCategory)
ggplot(grouped_effortCategory, aes(x = effortCategory, y = meanRT, fill=handType)) +
  geom_boxplot(fill=c("lightblue", "cornsilk", "lavenderblush2")) +
  labs(x = "Effort Category",
       y = "Mean Response Time",
       title = "Response Time by Effort Category")
ggplot(grouped_effortCategory, aes(x = effortCategory, y = meanAcc)) +
  geom_boxplot(fill=c("lightsalmon", "palegreen3", "orchid4")) +
  labs(x = "Effort Category",
       y = "Mean Accuracy",
       title = "Accuracy by Effort Category")
ggplot(grouped_effortCategory, aes(x = handType, y = meanRT)) +
  geom_boxplot(fill=c("lightblue", "cornsilk", "lavenderblush2")) +
  labs(x = "Handedness",
       y = "Mean Response Time",
       title = "Response Time by Hand Placement")
ggplot(grouped_effortCategory, aes(x = handType, y = meanAcc, fill=trialHalf)) +
  geom_boxplot()+
  fill=c("lightblue", "cornsilk", "lavenderblush2") +
  labs(x = "Handedness",
       y = "Mean Accuracy",
       title = "Accuracy by Hand Placement")